miniKanren with fair search strategies

### 2019-05-02 Paper/paper.tex
make file and notes
sec 1

%% scheme-list :
\usepackage{listings}
\usepackage{color}
\usepackage{textcomp}

@@

\section{introduction}

miniKanren programs, especially relational interpreters, have been proven to be 
useful in solving many problems \citep{byrd2017unified}. A subtlety in writing 
relational programs having large \conde expressions, such as interpreters, is 
that the order of \conde clauses can affect the speed considerably. When a 
\conde expression is large enough, the left clauses consume almost all the 
resource and the right ones are hardly explored. The unfair \disj{} of the 
current search strategy, interleaving DFS, is the cause. Under the hood, \conde 
uses \texttt{conj} to create a goal for each clause, and \texttt{disj} to 
combine these goals to one. The current \texttt{disj} allocates half resource 
to its first goal, then allocates the other half to the rest similarly, except 
for the last clause which takes all the resource.

Being aware of \texttt{disj} fairness, we also investigate \texttt{conj} 
fairness.

We propose three new search strategies, balanced interleaving DFS (DFS$_\textrm{bi}$), fair 
DFS (DFS$_\textrm{f}$), and BFS. They have different characteristics of fairness 
(Table.~\ref{fairness}). We prove that the pure subset of our BFS and the BFS 
proposed by Seres et
al~\citep{seres1999algebra} produce the same result when queried. But our code 
is shorter and runs faster. Because the two BFSs are equivalent, we do not 
distinguish them in most places.

\begin{table}
	\begin{tabular}{|c|c|c|c|c|}
		\hline 
		fairness & DFS$_\textrm{i}$ & DFS$_\textrm{bi}$ & DFS$_\textrm{f}$ & BFS \\ 
		\hline 
		disj & unfair & almost-fair & fair & fair \\ 
		\hline 
		conj & unfair & unfair & unfair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{fairness of search strategies}
	\label{fairness}
\end{table}

@@

In the following program, the three \conde clauses differ in a trivial way. So 
we expect lists of each sort constitute $1/4$ of the answer list. However, 
DFS$_\textrm{i}$, the current search strategy, gives us much more lists of \texttt{a} than 
other sorts of lists. And some sorts (e.g. lists of \texttt{c}) are hardly 
found. The situation would be worse if we add more \conde clauses.

@@

In the following program, the three \conde clauses differ in a trivial way. So 
we expect lists of each sort constitute $1/4$ of the answer list. However, 
search strategies with unfair \conj{} (e.g. DFS$_\textrm{i}$, DFS$_\textrm{f}$) give us much 
more lists of \texttt{a} than other sorts of lists. And some sorts (e.g. lists 
of \texttt{c}) are hardly found. Although DFS$_\textrm{i}$'s \disj{} is unfair in general, 
it is fair when there is no call to relational definition in sub-goals, 
including this case. The situation would be worse if we add more \conde 
clauses. The result with DFS$_\textrm{bi}$, whose \conj{} is also unfair, is similar, but 
due to its different \disj{}, the position of \texttt{b} and \texttt{c} are 
swapped.

>>>>

%% scheme-list :
\usepackage{listings}
\usepackage{color}
\usepackage{textcomp}
\usepackage{trackchanges}

\addeditor{MVC}

@@

\section{introduction}
\NOTE[MVC]{We need clear definitions of the following: fair, almost-fair, and unfair. Either use bullet points, or definition environment. Google ``latex math definition.''}
\NOTE[MVC]{A quick intro to miniKanren would be nice here.}

miniKanren programs, especially relational interpreters, have been proven to be 
useful in solving many problems \citep{byrd2017unified}. A subtlety in writing 
relational programs having \conde expressions with large number of clauses,
such as interpreters, is 
that the order of \conde clauses can affect the speed considerably. When a 
\conde expression is large enough, the left clauses consume almost all the 
resources and the right ones are barely explored. The cause is the unfair
search strategy, interleaving DFS(DFS$_\textrm{i}$). In a \conde expression,
each clause takes half of the received resources and pass the other half
to the remaining clauses, except for the last clause that takes all resources
it receives.

\NOTE[MVC]{A motivation is needed. Why do people want fairness?}
\NOTE[MVC]{\texttt{conj} fairness is not defined.}
Being aware of \texttt{disj} fairness, we also investigate \texttt{conj} 
fairness. 
We propose three new search strategies, balanced interleaving DFS (DFS$_\textrm{bi}$), fair 
DFS (DFS$_\textrm{f}$), and BFS. They have different characteristics of fairness 
(Fig.~\ref{fairness}). We prove that the pure subset of our BFS and the BFS 
proposed by Seres et
al~\citep{seres1999algebra} produce the same result when queried. But our code 
is simpler and runs faster on our benchmarks. Because the two BFSs are equivalent,
we do not distinguish them in most places.

\begin{figure}[tbp]
	\begin{tabular}{|c|c|c|c|c|}
		\hline 
		fairness & DFS$_\textrm{i}$ & DFS$_\textrm{bi}$ & DFS$_\textrm{f}$ & BFS \\ 
		\hline 
		disj & unfair & almost-fair & fair & fair \\ 
		\hline 
		conj & unfair & unfair & unfair & fair \\ 
		\hline 
	\end{tabular} 
	\caption{fairness of search strategies}
	\label{fairness}
\end{figure}

@@

In the following program, the three \conde clauses differ in a trivial way. So 
we expect lists of each sort constitute $1/4$ of the answer list. However, 
DFS$_\textrm{i}$, the current search strategy, gives us much more lists of \texttt{a} than 
other sorts of lists. And some sorts (e.g. lists of \texttt{c}) are rarely 
found. The situation would be worse if we add more \conde clauses.

@@

In the following program, the three \conde clauses differ in a trivial way. So 
we expect lists of each sort constitute $1/4$ of the answer list. However, 
search strategies with unfair \conj{} (e.g. DFS$_\textrm{i}$, DFS$_\textrm{f}$) give us much 
more lists of \texttt{a} than other sorts of lists. And some sorts (e.g. lists 
of \texttt{c}) are rarely found. Although DFS$_\textrm{i}$'s \disj{} is unfair in general, 
it is fair when there is no call to relational definition in sub-goals, 
including this case. The situation would be worse if we add more \conde 
clauses. The result with DFS$_\textrm{bi}$, whose \conj{} is also unfair, is similar, but 
due to its different \disj{}, the position of \texttt{b} and \texttt{c} are 
swapped.
